---
title: "Risky Business"
author: "Jack Carter"
date: "2/8/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(jsonlite)
library(knitr)
library(ggplot2)
library(ggthemes)

# reads in the risk headlines. 
data <- read_csv("C:\\Users\\HUAWEI\\Desktop\\Projects\\Risky-Business\\Data\\risk_headlines.csv") %>%
  filter(
    Collected >= 20220207
  )

# gets the z score for a specific set of data points.  
get_z_score <- function(data) {
  mean <- mean(
    data
  )
  sigma <- sd(
    data
  )
  z_scores <- list()
  for(i in 1:length(data)) {
    z_scores[i] <- (data[i] - mean) / sigma
  }
  return(unlist(
    z_scores
  ))
}

# get the z-scores for each category of a single continent. 
get_continent_z_scores <- function(continent) {
  data <- data[data["Continent"]==continent,]
  count <- data %>%
    group_by(
      Category
    ) %>%
    count()
  z_scores <- tibble(
    z_score = get_z_score(count$n)
  )
  final_data <- count[,1] %>%
    cbind(
      z_scores
    ) %>%
    mutate(
      Continent = continent
    )
  return(final_data)
}

# gets the z-scores for each category by continent.  
z_scores <- rbind_pages(
  lapply(
    unique(data$Continent), get_continent_z_scores
  )
)

# my personal plot theme for data visualizations. 
my_theme <- theme_economist_white(
  gray_bg = FALSE
) +
  theme(plot.title = element_text(
    hjust = 0.5, 
    vjust = 13, 
    size = 9, 
    color = "#474747"
  ),
  plot.margin = unit(
    c(1.5, 1, 1.5, 1), "cm"
  ),
  axis.text = element_text(
    size = 9, 
    color = "gray30"
  ),
  axis.text.x=element_text(
    vjust = -2.5
  ),
  axis.title.x = element_text(
    size = 9, 
    color = "gray30", 
    vjust = -10
  ),
  axis.title.y = element_text(
    size = 9, 
    color = "gray30", 
    vjust = 10
  ),
  legend.direction = "vertical", 
  legend.position = "right",
  legend.title = element_blank(),
  legend.text = element_text(
    size = 12, 
    color = "gray20"
  ),
  legend.margin=margin(
    1, -15, 1, 0
  ),
  legend.spacing.x = unit(
    0.25, "cm"
  ),
  legend.key.size = unit(
    1, "cm"
  ), 
  legend.key.height = unit(
    0.75, "cm"
  ),
  strip.text = element_text(
    hjust = 0.5, 
    vjust = 1, 
    size = 10, 
    color = "#474747"
  ),
  panel.spacing = unit(
    2, "lines")
  )

```

## **Summary**

This project uses web scraping, regex and a scoring mechanism to identify the different kinds of political violence risks that occur in the emerging world, as well as how their prevalence differs by continent and region.

&nbsp;

## **Method**

**1) Scraping:**

The websites of over 250 local newspapers in emerging regions are scraped three times a week, beginning February 7 2022. 

**2) Regex:**

A series of regex strings, including keywords, actor/action combinations and hyperbolic situation scenarios, are used to create multiple word banks.

**3) Scoring Mechanism:**

The word banks are used to score and label the headlines as one of 5 kinds of political violence risks. 

**4) Z-scores:**

The relative distribution of risks is then compared using a standardized score according to the relevant continent or region. 

&nbsp;

## Results

### **1) Overview:**

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# overview. 
overall_data <- data %>%
  group_by(Category, Continent) %>%
  count()

overall_data %>%
  ggplot(
    aes(x=Continent, y=n, fill=Category)
  ) +
  geom_bar(
    stat = "identity", position = "fill"
  ) +
  ggtitle(
    "Continent"
  ) +
  ylab(
    "Relative Distribution"
  ) +
  xlab(
    ""
  ) + 
  my_theme +
  theme(
    axis.text.y=element_blank()
  ) +
  scale_fill_discrete(
    name = "Risks", labels = c("Military Conflict", 
                               "Organized Crime", 
                               "Regime Instability",
                               "Social Unrest",
                               "Terrorism")
    ) +
  scale_x_discrete(
    guide = guide_axis(n.dodge=2)
    )
```